$p$-adic precision

### 2014-02-05 padicprec.tex
Conflict resolved

\begin{proof}
We may assume that $v_0=0$ and $f(0)=0$. Since $f'(0)$ is surjective, by 
the open mapping Theorem, there exists $C>0$ such that $B_F(1) \subset 
C \cdot f'(0) \cdot B_E(1)$. Let $\varepsilon>0$ be such that 
$\varepsilon C < 1$. Let $U_\varepsilon \subset E$ be given by the fact 
that $f$ is differentiable at $0$. We may assume $U_\varepsilon = 
B_E(\delta)$ for some $\delta >0$.

Let $r \in (0, \delta)$.
Let us show that $f$ maps surjectively $H$ onto $f'(0) \cdot H$.
Let $x \in H$. We must show that $f(x) \in f'(0) \cdot H$. 
By differentiability at $0$, $\Vert f(x)-f'(0) \cdot x \Vert \leq 
\varepsilon \Vert x \Vert $. Therefore, there exists $y \in f'(0) \cdot 
H$ such that 
$\Vert y \Vert \leq \varepsilon r$. By the open mapping Theorem and the 
surjectivityy of $f'(0)$, there exists $x_1 \in E$ such that $f'(0) \cdot 
x_1 =y$ and $\Vert x_1 \Vert \leq \varepsilon C \cdot r < r$. Thus, $x_1 
\in B^-_E(r) \subset H$ and $f(x)= f'(0) \cdot (x-x_1) \in f'(0) 
\cdot H$.

We shall now prove surjectivity. Let $y \in f'(0) \cdot H$. 
Let $x_0 \in H$ be such that $y = f'(0) \cdot x_0$.
We define by induction two sequences $(x_n)$ and $(z_n)$ by
\begin{itemize}
\item $z_n \in E$ is such that
$f'(0) \cdot z_n = y - f(x_n)$ and
$\Vert z_n \Vert \leq C \cdot \Vert y - f(x_n) \Vert$,
\item $x_{n+1}=x_n+z_n$.
\end{itemize}
By definition, $\Vert z_0 \Vert < C \Vert y_0\Vert$ and $y_0 
= -f(x_0)+f'(0) \cdot x_0$. Therefore, $\Vert y_0 \Vert < \varepsilon 
\Vert x_0 \Vert \leq \varepsilon r$. Thus $\Vert z_0 \Vert \leq 
\varepsilon C \cdot r < r$. Thus, $x_1$ and $z_0$ lie in $B^-_E(r) 
\subset H$. An easy induction involving an analogous computation
shows that the sequences $(x_n)$ and $(z_n)$ are well defined and
take their values in $H$.

Now, we notice that
$y - f(x_{n+1}) = f(x_n) - f(x_{n+1}) + f'(0) \cdot z_n$.
Therefore, using differentiability, we get:
$$\Vert y  - f(x_{n+1}) \Vert 
\leq \varepsilon \Vert z_n \Vert \leq \varepsilon C \cdot \Vert 
y - f(x_n) \Vert,$$
from what we deduce that $\Vert y - f(x_{n+1}) \Vert = 
O(a^n)$ and $\Vert z_n \Vert = O(a^n)$ with $a = \varepsilon C < 1$.
This shows that $(x_n)$ is a Cauchy sequence and therefore converges 
(since $E$ is complete). If $x$ denotes the limit of $(x_n)$, we have
$x \in H$ and $y=f(x)$. The result is proven.
\end{proof}

@@

Before discussing applications of Lemma \ref{lem:main} to ultrametric 
precision, we would like to describe a little bit more restrictive 
setting where the constant $\delta$ appearing in this Lemma can be 
computed explicitly.
We refer to Schneider \cite{Schneider}, \S 6, for a definition of a 
locally analytic function.

\begin{rmk}
Let $f$ be locally analytic function $K^n \rightarrow K^m$. Let $x \in K^n$ be such that $f'(x)$ is surjective. Then the radius of the ball needed can be read from the Newton polygon given by the norms of the Taylor series expansion of $f$ at $x$. We assume the $C$ given by the Banach-Schauder is given, $\varepsilon = \frac{1}{C}$ and now, we try to recover the link between $\varepsilon$ and $\delta$.
\end{rmk}

\begin{prop}
\label{prop:locanalytic}
\todo{Write a proposition.}
\end{prop}

\begin{proof}
Let us assume that $x=0$, $f(0)=0$, and for $r>0$, $f(x)=f'(0) \cdot x +\sum_{n \geq 2} \frac{f^{(n)}(0)}{n!} \cdot (x, \dots,x)$ is the Taylor series expansion of $f$ at $0$, coinciding on $B_E (0,r)$.

We shall now find $l$ such that if $x \in B_{K^n}(0,l)$, then $\Vert \sum_{n \geq 2} \frac{f^{(n)}(0)}{n!}(x,\dots,x) \Vert \leq \varepsilon$.

By continuity and ultrametricity, it is enough for $l$ to be such that, if $x \in B_{K^n}(0,l)$, then  \[ \sup_{n \geq 2} \interleave f^{(n)}(0) \interleave \frac{\Vert x \Vert^n }{ \vert n! \vert } \leq \varepsilon \Vert x \Vert .\]

Therefore, \[ \sup_{n \geq 2} \interleave f^{(n)}(0) \interleave \frac{l^{n-1} }{\vert n! \vert } \leq \varepsilon . \]

In other words, $\forall n \geq 2$, $\interleave f^{(n)}(0) \interleave \frac{l^{n-1} }{\vert n! \vert } \leq \varepsilon.$

By passing to logarithm and valuation, we get for all $n \geq 2$: \[ \log_p (\interleave f^{(n)}(0) \interleave )-(n-1) \log_p l-val (n!)\geq - \log_p \varepsilon.\]

By Legendre formula, it is enough that for all $n \geq 2$:
\[
\log_p (\interleave f^{(n)}(0) \interleave )-(n-1) (\log_p l-(n-1)) \geq - \log_p \varepsilon.
\]

If we denote by $NP_F$ the Newton polygon application of the power series 
\[
F(t) = \sum_{k\geq 0}^{+\infty} \log_p (\interleave f^{(k+1)}(0) \interleave ) t^k
\]
then it is enough that:
\[
\sup_{ x \in \mathbb{R}_+} \left( NP_F(x) -x  \left( \log_p l - \frac{1}{p-1} \right) \right) \geq - \log_p \varepsilon.
\]

If $NP_F^*$ is the Legendre-Fanchel transform of $NP_F$, then it is enough that:
\[ NP_F^* \left( \log_p l - \frac{1}{p-1} \right) \leq \log_p \varepsilon.\]

Therefore, by definition of the Legendre-Fanchel transform, it is enough to:
\begin{itemize}
\item take $M=(0,\log_p \varepsilon)$,
\item draw the line passing by $M$ tangent to the graph of $NP_F$,
\item if we denote by $\mu$ the slope of this line, then take $l$ such that $\mu + \frac{1}{p-1}=\log_p l$.
\end{itemize}
\end{proof}

>>>>

\begin{proof}
We may assume that $v_0=0$ and $f(0)=0$. Since $f'(0)$ is surjective, by 
the open mapping Theorem, there exists $C>0$ such that $B_F(1) \subset 
C  f'(0) \cdot B_E(1)$. Let $\varepsilon>0$ be such that 
$\varepsilon C < 1$. Let $U_\varepsilon \subset E$ be given by by definition \ref{deftn : diff} and the fact that $f$ is differentiable at $0$. We may assume $U_\varepsilon = 
B_E(\delta)$ for some $\delta >0$.

Let $r \in (0, \delta)$. Let $B^-_E(r) \subset H \subset
B_E(r).$
Let us show that $f$ maps surjectively $H$ onto $f'(0) \cdot H$.

We first prove that $f(H) \subset f'(0) \cdot H$.
Let $x \in H$. By differentiability at $0$, $\Vert f(x)-f'(0) \cdot x \Vert \leq 
\varepsilon \Vert x \Vert $. Therefore, if $y=f(x)-f'(0) \cdot x,$ then $\Vert y \Vert \leq \varepsilon r$. As a consequence of the open mapping Theorem, $B_F(\varepsilon r) \subset 
\varepsilon C  f'(0) \cdot B_E(r),$ and there exists $x_1 \in \varepsilon C  \cdot B_E(r)$ such that $f'(0) \cdot x_1 =y.$ Since $\varepsilon C <1,$  $x_1 \in B^-_E(r) \subset H.$ and finally, $f(x)= f'(0) \cdot (x-x_1) \in f'(0) \cdot H$.

We shall now prove surjectivity. Let $y \in f'(0) \cdot H$. 
Let $x_0 \in H$ be such that $y = f'(0) \cdot x_0$. Since $f(x_0) \in f'(0) \cdot H,$ we can define $z_0 \in E$ to be such that $f'(0) \cdot z_0 = y - f(x_0)$ and $\Vert z_0 \Vert \leq C \cdot \Vert y - f(x_0) \Vert$.
We then define by induction two sequences $(x_n)$ and $(z_n)$ by
\begin{itemize}
\item $x_{n+1}=x_n+z_n$ and 
\item $z_n \in E$ is such that
$f'(0) \cdot z_n = y - f(x_n)$ and
$\Vert z_n \Vert \leq C \cdot \Vert y - f(x_n) \Vert$.
\end{itemize}
By definition, $\Vert z_0 \Vert < C \Vert y_0\Vert$ and $y 
= -f(x_0)+f'(0) \cdot x_0$. Therefore, $\Vert y \Vert < \varepsilon 
\Vert x_0 \Vert \leq \varepsilon r$. Thus $\Vert z_0 \Vert \leq 
\varepsilon C \cdot r < r$. Thus, $z_0$ and $x_1=x_0+z_0$  lie in $B^-_E(r) 
\subset H$. An easy induction involving an analoguous computation
shows that the sequences $(x_n)$ and $(z_n)$ are well defined and
take their values in $H$.

We notice that $y - f(x_{n+1}) = f(x_n) - f(x_{n+1}) + f'(0) \cdot z_n$.
Since $z_n=x_{n+1}-x_n,$, if we use differentiability, we get:
$$\Vert y  - f(x_{n+1}) \Vert 
\leq \varepsilon \Vert z_n \Vert \leq \varepsilon C \cdot \Vert 
y - f(x_n) \Vert,$$
from what we deduce that $\Vert y - f(x_{n+1}) \Vert = 
O(a^n)$ and $\Vert z_n \Vert = O(a^n)$ with $a = \varepsilon C < 1$.
This shows that $(x_n)$ is a Cauchy sequence and, since $E$ is complete, it converges. Because $H$ is closed, if $x$ denotes the limit of $(x_n)$, we have $x \in H.$ Yet, as a consequence of the differentiability, $f$ is continuous over $U_\varepsilon.$ Therefore, since $x \in H \subset U_\varepsilon,$ then  $y=f(x)$. The result is proven.
\end{proof}

@@

Before discussing applications of Lemma \ref{lem:main} to ultrametric 
precision, we would like to describe a little bit more restrictive 
setting where the constant $\delta$ appearing in this Lemma can be 
computed explicitly.

Indeed, when $f$ is a locally analytic function, we prove that we can recover $\delta$ from a Newton polygon defined with the Taylor series expansion of $f.$

We refer to Schneider \cite{Schneider}, \S 6, for a definition of a 
locally analytic function.

\begin{prop}\label{prop:locanalytic}
Let $E$ and $F$ be two complete normed $K$-vector spaces, and let $f$ be locally analytic map $E \rightarrow F$.  Let $x \in E$, and let $f(x+h)=f(x)+\sum_{k \geq 0}^{+\infty} L_k \cdot (h,\dots,h)$ be the power series development of $f$ at $x$. We assume $L_1$ is surjective. 
Let $C$ be given by the Banach-Schauder theorem for $L_1$, $\varepsilon = \frac{1}{C}$. 
Let $F(t) = \sum_{k\geq 0}^{+\infty} \interleave L^{k+1}(0) \interleave  t^k \in \mathbb{R}[[t]]$, let $NP_F$ be its Newton polygon.
Let $\mu$ be the slope of the line passing by $(0, \log_p \varepsilon)$ and tangent to $NP_F$.
Then $r$ such that $\mu=\log_p r$ is enough for lemma \ref{lem:main}.
\end{prop}
\begin{proof}
Let us assume that $x=0$, $f(0)=0$, and for $r>0$, $f(x)=L_1 \cdot x +\sum_{n \geq 2} L_n \cdot (x, \dots,x)$ is the Taylor serie expansion of $f$ at $0$, coinciding on $B_E (0,r)$.

We shall now find $l$ such that if $x \in B_{K^n}(0,l)$, then $\Vert \sum_{n \geq 2} L_n \cdot (x,\dots,x) \Vert \leq \varepsilon \vert x \vert $.

By continuity and ultrametricity, it is enough for $l$ to be such that, if $x \in B_{K^n}(0,l)$, then  \[ \sup_{n \geq 2} \interleave L_n \interleave \Vert x \Vert^n \leq \varepsilon \Vert x \Vert .\]

Therefore, \[ \sup_{n \geq 2} \interleave L_n \interleave l^{n-1}  \leq \varepsilon . \]

In other words, $\forall n \geq 2$, $\interleave L_n \interleave \l^{n-1}  \leq \varepsilon.$

By passing to logarithm, we get for all $n \geq 2$ : \[ \log_p (\interleave L_n \interleave )-(n-1) \log_p l \geq - \log_p \varepsilon.\]

If we denote by $NP_F$ the Newton polygon application of the power serie $F(t) = \sum_{k\geq 0}^{+\infty} \interleave L_{k+1} \interleave  t^k$, then it is enough that :
\[ \sup_{ x \in \mathbb{R}_+} \left( NP_F(x) -x  \log_p l \right) \geq - \log_p \varepsilon.\]

If $NP_F^*$ is the Legendre-Fanchel transform of $NP_F$, then it is enough that :
\[ NP_F^* \left( \log_p l  \right) \leq \log_p \varepsilon.\]

Therefore, by definition of the Legendre-Fanchel transform, it is enough to :
\begin{itemize}
\item take $M=(0,\log_p \varepsilon)$,
\item draw the line passing by $M$ tangent to the graph of $NP_F$ (it could be tangent at infinity),
\item if we denote by $\mu$ the slope of this line, then take $l$ such that $\mu=\log_p l$.
\end{itemize}
\end{proof}

We remark that when considering finite-dimensional applications over $p$-adic fields, a more refined proposition is available.

\begin{prop} \label{prop:locanalyticfinitedim}
If $f$ is a locally analytic map $K^n \rightarrow K^m$, with $K$ a $p$-adic field, if $C$ and $\varepsilon$ are as above. It is enough to work with the Newton polygon $NP_F$ of $F(t) = \sum_{k\geq 0}^{+\infty} \interleave f^{(k+1)}(0) \interleave  t^k \in \mathbb{R}[[t]]$ : Let $\mu$ be the slope of the line passing by $(0, \log_p \varepsilon)$ and tangent to $NP_F$.
Then $r$ such that $\mu+\frac{1}{p-1}=\log_p r$ is enough for lemma \ref{lem:main}. 
\end{prop}
\begin{proof}
In comparison with the previous proposition, $L_n=\frac{f^{(n)}(0)}{n!}$. Therefore, by Legendre formula, it would be enough that for all $n \geq 2$ :
\[\log_p (\interleave f^{(n)}(0) \interleave )-(n-1) (\log_p l-(n-1)) \geq - \log_p \varepsilon,\]
and therefore, \[ NP_F^* \left( \log_p l - \frac{1}{p-1} \right) \leq \log_p \varepsilon.\]
The result follows \textit{mutatis mutandis}.
\end{proof}
